---
title: "Problem 5"
author: "Nidhi"
date: "2022-11-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 5: Computational Finance - Modelling Stock prices

1. Estimate the parameters of the models $\theta=(\alpha,\beta,\sigma)$ using the method of moments type plug-in estimator discussed in the class.

2. Estimate the parameters using the `lm` built-in function of `R`. Note that `lm` using the OLS method.

3. Fill-up the following table

Parameters | Method of Moments | OLS
-----------|-------------------|-----
$\alpha$   |                   |
$\beta$    |                   |
$\sigma$   |                   |

4. If the current value of Nifty is 18000 and it goes up to 18200. The current value of TCS is Rs. 3200/-. How much you can expect TCS price to go up?

```{r, warning=FALSE}
library(quantmod)
getSymbols('TCS.NS')
getSymbols('^NSEI')


TCS_rt <- diff(log(TCS.NS$TCS.NS.Adjusted))
Nifty_rt <- diff(log(NSEI$NSEI.Adjusted))

retrn <- cbind.xts(TCS_rt, Nifty_rt)
retrn <- na.omit(data.frame(retrn))

names(retrn) = c('TCS', 'Nifty50')

covariance <- cov(retrn[,'Nifty50'], retrn[, 'TCS'])
nifty.var <- var(retrn[, 'Nifty50'])
beta <- covariance/nifty.var

tcs.mean <- mean(retrn[,'TCS'])
nifty.mean <- mean(retrn[, 'Nifty50'])
alpha <- tcs.mean - beta*nifty.mean

sigma = sqrt((sum((retrn[,'TCS']-alpha-beta*retrn[,'Nifty50'])^2))/length(retrn[,'TCS']))

fit <- summary(lm(TCS ~ Nifty50, data = retrn))

ols.alpha <- fit$coefficients[1]
ols.beta <- fit$coefficients[2]
ols.sigma <- fit$sigma

TCS_return = alpha + beta*log(18200/18000)
TCS_new = 3200*exp(TCS_return)

```

Parameters | Method of Moments | OLS
-----------|-------------------|-----
$\alpha$   |0.0004616524       |0.0004616524
$\beta$    |0.7436617          |0.7436617
$\sigma$   |0.01618042         |0.01618481
