---
title: "Probability and Statistics in R"
author: "Assignment 2 Problem 2"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

## Problem 2 : Simulation Study to Understand Sampling Distribution

**Part A** Suppose
$X_1,X_2,\cdots,X_n\stackrel{iid}{\sim} Gamma(\alpha,\sigma)$, with pdf
as $$
f(x | \alpha,\sigma)=\frac{1}{\sigma^{\alpha}\Gamma(\alpha)}e^{- x/\sigma}x^{\alpha-1},~~~~0<x<\infty,
$$ The mean and variance are $E(X)=\alpha\sigma$ and
$Var(X)=\alpha\sigma^2$. Note that `shape =` $\alpha$ and `scale =`
$\sigma$.

1.  Write a `function` in `R` which will compute the MLE of
    $\theta=\log(\alpha)$ using `optim` function in `R`. You can name it
    `MyMLE`
2.  Choose `n=20`, and `alpha=1.5` and `sigma=2.2`
    (i) Simulate $\{X_1,X_2,\cdots,X_n\}$ from
        `rgamma(n=20,shape=1.5,scale=2.2)`
    (ii) Apply the `MyMLE` to estimate $\theta$ and append the value in
         a vector
    (iii) Repeat the step (i) and (ii) 1000 times
    (iv) Draw histogram of the estimated MLEs of $\theta$.
    (v) Draw a vertical line using `abline` function at the true value
        of $\theta$.
    (vi) Use `quantile` function on estimated $\theta$'s to find the 2.5
         and 97.5-percentile points.
3.  Choose `n=40`, and `alpha=1.5` and repeat the (2).
4.  Choose `n=100`, and `alpha=1.5` and repeat the (2).
5.  Check if the gap between 2.5 and 97.5-percentile points are
    shrinking as sample size `n` is increasing?

```{r, warning=FALSE, message=FALSE}

simulation <- function(n, alpha, sigma, N){
  
  theta = log(alpha)
  
  logl <- function(par, sample){
    loglike = sum(dgamma(sample, shape = par[1], scale = par[2], log = T))
    return(-loglike)
  }

  mymles = c()
  for(i in 1:N){
    sample = rgamma(n = n, shape = alpha, scale = sigma)
    par = c(1.5, 2.2)
    fit = optim(par, logl, sample = sample)
    alpha_hat = fit$par[1]
    mymle = log(alpha_hat)
    mymles = c(mymles, mymle)
  }
  
  hist(mymles, col = "skyblue", xlab = "MLE", 
       main = "Histogram of estimated MLEs of theta")
  abline(v = theta, col = "red")
  quantiles = quantile(mymles, probs = c(0.025, 0.975))
  print("Quantiles")
  print.table(quantiles)
  diffq = quantiles[2] - quantiles[1]
  print(paste("Gap between 2.5 and 97.5-percentile points", diffq))
  
}

```

## For sample size:20
```{r}
simulation(20, 1.5, 2.2, 1000)
```
\newpage

## For sample size:40
```{r}
simulation(40, 1.5, 2.2, 1000)
```
\newpage

## For sample size:80
```{r}
simulation(80, 1.5, 2.2, 1000)
```
\newpage

## For sample size:100 and simulation size:1000
```{r}
simulation(100, 1.5, 2.2, 1000)
```
\newpage

## For sample size:100 and simulation size:2000
```{r}
simulation(100, 1.5, 2.2, 2000)
```
\newpage

## For sample size:200 and simulation size:1000
```{r}
simulation(200, 1.5, 2.2, 1000)
```
\newpage

## For sample size:200 and simulation size:2000
```{r}
simulation(200, 1.5, 2.2, 2000)
```

## From the above results we can observe that as the sample size n is increasing, the gap between 2.5 and 97.5 percentile points is shrinking.

